Notes on Perturbational Quantities

### 2008-07-29 docs/pert_notes/paper.tex
more work on rho and rhoh sections

superscript indicates that we have not called correct\_base yet.  We compute $\rho$ on edges 
by adding $\rho_0^{n+\half} = (\rho_0^n + \rho_0^{(2)}) / 2$ to $\rho'$ at edges.  We use 
$\rho^{(2)}$ rather than $\rho^{n+1}$ to be consistent with the fact that we omitted the 
$\eta_\rho$ contribution to $\rho'$ at edges.  For the non-radial faces, we can directly add 
$\rho_0^{n+\half}$ to $\rho'$ since it is a cell-centered quantity.  For the radial faces, we 
interpolate to obtain $\rho_0^{n+\half}$ at edges.

@@

\subsection{Predicting $\rho_0$}
To compute $\eta_\rho$, we need $\rho_0$ at edge-centered, time-centered locations.  
However, we do NOT use the same edge-centered, time-centered $\rho_0^{n+\half}$ that was 
referred to in Section \ref{Computing rho at edges}.  Instead, we use a $\rho_0$ computed 
with a Godunov method in make\_edge\_state\_1d.  We call make\_edge\_state\_1d using the 
term proportional to $\rho_0$ in  (\ref{rho0 equation}) as the forcing.  The $\eta_\rho$ 
term is not 
included.\MarginPar{I think there is a bug in the code and the $\eta_\rho$ term needs to 
be included here.}  This gives us the so-called $\rho_0^{n+\half,{\rm pred}}$.

@@

Full enthalpy equation:

@@

+ \underbrace{\psi + \Ubt\cdot\nabla p_0}_{\frac{Dp_0}{Dt}} + \nabla\cdot\kappa\nabla T 
+ \rho H_{\rm nuc} + \rho H_{\rm ext}\label{rhoh equation}

@@

+ (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k 

@@

To simplify matters, the we have already called react\_state, so for now we will ignore 
the $\rho H_{\rm nuc}$ and $\rho H_{\rm ext}$ terms in the remaining derivations.  
Subtracting (\ref{rhoh0 equation}) from (\ref{rhoh equation}) and ignoring the 
aforementioned terms gives the perturbational enthalpy equation:
\begin{eqnarray}
\frac{\partial(\rho h)'}{\partial t} &=& -\nabla\cdot[(\rho h)'\Ub] 
- \nabla\cdot[(\rho h)_0\Ubt] + \wt\frac{\partial p_0}{\partial r} 
+ (\nabla\cdot\kappa\nabla T - \overline{\nabla\cdot\kappa\nabla T}) \nonumber \\
&=& -\Ub\cdot\nabla(\rho h)' \underbrace{- (\rho h)'\nabla\cdot\Ub 
- \nabla\cdot[(\rho h)_0\Ubt]} + \wt\frac{\partial p_0}{\partial r} 
+ (\nabla\cdot\kappa\nabla T 
- \overline{\nabla\cdot\kappa\nabla T})\label{rhohprime equation}
\end{eqnarray}
We use make\_edge\_scal to get $(\rho h)'$ at edges.  The function modify\_scal\_force 
computes the underbraced term in (\ref{rhohprime equation}).  The function mkrhohforce 
computes the $\wt$ term.\MarginPar{I need to write a function to handle the thermal 
diffusion terms.}
\subsubsection{Predicting $(\rho h)_0$ at edges}\label{Predicting rhoh0 at edges}
After the call to react\_state, we set $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$.  
We use the Godunov algorithm in make\_edge\_state\_1d to predict $(\rho h)_0$ at edges, 
omitting the terms added in react\_state, so we are essentially evolving an abbreviated 
version of (\ref{rhoh0 equation}):
\begin{eqnarray}
\frac{\partial(\rho h)_0}{\partial t} &=& -\nabla\cdot[(\rho h)_0 w_0\eb_r] + \psi 
+ \overline{\nabla\cdot\kappa\nabla T} \nonumber \\
&=& -w_0\frac{\partial(\rho h)_0}{\partial r} - (\rho h)_0\frac{\partial w_0}{\partial r} 
+ \psi + \overline{\nabla\cdot\kappa\nabla T}.\label{rhoh0 equation}
\end{eqnarray}
In advect\_base, the forcing term is set to the term proportional to 
$(\rho h)_0$.\MarginPar{We are currently omitting the $\psi$ and thermal conduction 
terms.  I think this is a bug.}
\subsubsection{Computing $\rho h$ at edges}
Simply add the results from Sections \ref{Predicting rhohprime at edges} and 
\ref{Predicting rhoh0 at edges}, using spatial interpolation to get $(\rho h)_0$ 
at non-radial faces.
\subsection{Predicting $h$ (enthalpy\_pred\_type = 2)}
\subsection{Predicting $T$ into $(\rho h)'$ (enthalpy\_pred\_type = 3)}
\subsection{Predicting $T$ into $h$ (enthalpy\_pred\_type = 4)}
\end{document}

>>>>

superscript indicates that we have not called correct\_base yet (correct\_base takes 
$\rho_0^{(2)}$ as an input and computes $\rho^{n+1}$ as the output).  We compute $\rho$ on 
edges by adding $\rho_0^{n+\half} = (\rho_0^n + \rho_0^{(2)}) / 2$ to $\rho'$ at edges.  We 
use $\rho^{(2)}$ rather than $\rho^{n+1}$ to be consistent with the fact that we did not 
include an $\eta_\rho$ contribution to compute $\rho'$ at edges.  For the non-radial edges, 
we can directly add $\rho_0^{n+\half}$ to $\rho'$ since $\rho_0^{n+\half}$ is a cell-centered 
quantity.  For the radial edges, we interpolate to obtain $\rho_0^{n+\half}$ at radial edges 
before adding it to $\rho'$.

@@

\subsection{Predicting $\rho_0$ (predict\_rho = T or F)}
To advect $\rho_0$ in advect\_base and to compute $\eta_\rho$ that will be used in 
correct\_base, we need $\rho_0$ at edge-centered, time-centered locations.  However, we do 
not use the same edge-centered, time-centered $\rho_0^{n+\half}$ referred to in Section 
\ref{Computing rho at edges}.  Instead, we use a $\rho_0$ computed with a Godunov method in 
make\_edge\_state\_1d.  We call make\_edge\_state\_1d using the term proportional to 
$\rho_0$ in (\ref{rho0 equation}) as the forcing.  This gives us $\rho_0^{n+\half,{\rm pred}}$.
\MarginPar{You could argue that including a time-lagged $\eta_\rho$ source term would
increase the order of accuracy, but Ann has convinced me that this term isn't necessary.}

@@

The full enthalpy equation is

@@

+ \underbrace{\psi + \wt\frac{\partial p_0}{\partial r}}_{\frac{Dp_0}{Dt}} 
+ \nabla\cdot\kappa\nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext}. \label{rhoh equation}

@@

+ \wt\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k 

@@

To simplify matters, the we have already called react\_state, so for now we will ignore 
the $\rho H_{\rm nuc}$ and $\rho H_{\rm ext}$ terms in the remaining derivations.  
Subtracting (\ref{rhoh0 equation}) from (\ref{rhoh equation}) and ignoring the 
aforementioned terms gives the perturbational enthalpy equation:
\begin{eqnarray}
\frac{\partial(\rho h)'}{\partial t} &=& -\nabla\cdot[(\rho h)'\Ub] 
- \nabla\cdot[(\rho h)_0\Ubt] + \wt\frac{\partial p_0}{\partial r} 
+ (\nabla\cdot\kappa\nabla T - \overline{\nabla\cdot\kappa\nabla T}) \nonumber \\
&=& -\Ub\cdot\nabla(\rho h)' \underbrace{- (\rho h)'\nabla\cdot\Ub 
- \nabla\cdot[(\rho h)_0\Ubt]} + \wt\frac{\partial p_0}{\partial r} 
+ (\nabla\cdot\kappa\nabla T 
- \overline{\nabla\cdot\kappa\nabla T})\label{rhohprime equation}
\end{eqnarray}
We use make\_edge\_scal to get $(\rho h)'$ at edges.  The function modify\_scal\_force 
computes the underbraced term in (\ref{rhohprime equation}).  The function mkrhohforce 
computes the $\wt$ term.\MarginPar{I need to write a function to handle the thermal 
diffusion terms.}
\subsubsection{Predicting $(\rho h)_0$ at edges}\label{Predicting rhoh0 at edges}
After the call to react\_state, we set $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$.  
We use the Godunov algorithm in make\_edge\_state\_1d to predict $(\rho h)_0$ at edges, 
omitting the terms added in react\_state, so we are essentially evolving an abbreviated 
version of (\ref{rhoh0 equation}):
\begin{eqnarray}
\frac{\partial(\rho h)_0}{\partial t} &=& -\nabla\cdot[(\rho h)_0 w_0\eb_r] + \psi 
+ \overline{\nabla\cdot\kappa\nabla T} \nonumber \\
&=& -w_0\frac{\partial(\rho h)_0}{\partial r} - (\rho h)_0\frac{\partial w_0}{\partial r} 
+ \psi + \overline{\nabla\cdot\kappa\nabla T}.\label{rhoh0 equation}
\end{eqnarray}
In advect\_base, the forcing term is set to the term proportional to 
$(\rho h)_0$.\MarginPar{We are currently omitting the $\psi$ and thermal conduction 
terms.  I think this is a bug.}
\subsubsection{Computing $\rho h$ at edges}
Simply add the results from Sections \ref{Predicting rhohprime at edges} and 
\ref{Predicting rhoh0 at edges}, using spatial interpolation to get $(\rho h)_0$ 
at non-radial edges.
\subsection{Predicting $h$ (enthalpy\_pred\_type = 2)}
\subsection{Predicting $T$ into $(\rho h)'$ (enthalpy\_pred\_type = 3)}
\subsection{Predicting $T$ into $h$ (enthalpy\_pred\_type = 4)}
\end{document}
